Processing of visual and non-visual naturalistic spatial information in the "parahippocampal place area"

### 2021-02-15 paper/main.tex
intro: honing & typos

% brain mapping via fMRI
Humans can perceive and make sense of their daily environment casually and
seemingly without any effort.
%
Studies in the field of neuropsychology and neuroimaging (e.g.
\citep{penfield1950cerebral, fox1984noninvasive}) have shown that different
parts of the brain are specialized for different perceptual and cognitive
functions.
% occipital cortex for vision -> two pathways
The occipital cortex is considered to be primarily involved in the early stages
of visual perception and giving rise to two distinct pathways that process
different kind of visual information:
% dorsal vs ventral pathway
a dorsal stream (the ``where pathway'') that leads into the parietal lob , and a
ventral stream (the ``what pathway'') that leads into into the temporal lobe
\citep{goodale1992separate, mishkin1982contribution}.
% PPA
A classic example of a higher-level visual area in the ventral pathway is the
``parahippocampal place area'' (PPA) \citep{epstein1998ppa,
epstein1999parahippocampal}.
% anatomical location != functional location
The PPA is located in the posterior parahippocampal gyrus including adjacent
regions of the fusiform gyrus and anterior lingual gyrus
\citep{epstein2008parahippocampal}:
% neural correlate of scene perception
increased hemodynamic activity is observed in the PPA when subjects view
pictures of landscapes, buildings or landmarks (compared to e.g. pictures of
tools or faces) during blood oxygenation level-dependent functional magnetic
resonance imaging (BOLD fMRI) \citep{aguirre1998area, epstein2014neural,
epstein1998ppa, troiani2012object}.

% literature review overview 1: imagination & haptic exploration
Increased hemodynamic activity in the PPA generalizes to mental imagination of
landscapes \citep{ocraven2000mental} and haptic exploration of scenes
constructed from LEGO blocks \citep{wolbers2011modality}.
% O'Craven: watching pictures
In a study conducted by \cite{ocraven2000mental} participants viewed alternating
blocks of pictures showing famous faces and familiar places during an initial
experimental paradigm.
% O'Craven: mental imagery
In a subsequent paradigm, participants were instructed to ``form a vivid mental
image'' of the previously viewed pictures.
% O'Craven results
The PPA showed increased activation during imagination of places compared to
faces but the imagination tasks showed a smaller activation level compared to
the perceptual task.
% Wolbers: haptic exploration
In a block design study conducted by \cite{wolbers2011modality} the PPA of
[sighted] participants showed increased activation during a delayed
match-to-sample task of haptically explored scenes constructed from LEGO bricks
compared to haptically explored abstract geometric objects.
% Wolbers: connectivity analysis

% Aziz (2008): place related sentences
To our knowledge only one study \citep{aziz2008modulation} compared hemodynamic
activity levels in the PPA that were correlated with different semantic
categories occurring in spoken language.
% Aziz' stimuli; ``The Taj Mahal faces a long thin reflecting pool'', ``Marilyn
% Monroe has a large square jaw'', ``The television has a long antenna''
\cite{aziz2008modulation} used sentences that described famous or generic places
and faces, or objects.
% Aziz' tasks:
Participants were instructed to press a button whenever the sentence described
an inaccurate or improbable fact (e.g. ``Marilyn Monroe has a large square
jaw'')
% Aziz' results:
Activation in the left, but not right, PPA was significantly reduced when
participants listened to place-related sentences compared to listening to
face-related sentences. Moreover, this effect was only observed in sentences
involving famous places.

% summary if literature review
In summary, studies suggest that the PPA does preferably but not exclusively
respond to visually presented spatial information.
% three things in common
All reviewed studies share three common aspects:
% designed stimulus set
First, they employed a small sets of carefully chosen and conceptualized
stimuli.
% block-design & task
Second, studies used block-design paradigms to maximize detection power.
Further, they usually employed an explicit (perceptual) judgement task.
% conceptualized stimuli and block: disadvantages
Block-design studies that use conceptualized stimuli and a task have the
advantage of controlling confounding variables (e.g. color, luminance, size,
spatial frequencies, sentence length), maximizing detection power, and keeping
participants paying attention to the stimuli.
% conceptualized stimuli and block: disadvantages
Nevertheless, small sets of conceptualized stimuli and block-design paradigms
lack external and ecological validity \citep{westfall2016fixing,
hasson2004intersubject} because they poorly resemble how we, free of a[n
explicit] perceptual task, experience our rich/multidimensional and continuous
environment that our brains are accommodated to
\citep{sonkusare2019naturalistic}.
% Moreover, study participants can and will become aware about which perceptual
% process is examined in the study by merely being exposed to the experimental
% paradigm.
Lastly, the reviewed studies averaged results across 7-11 subjects.
% averaging pros and cons
The averaging approach improves signal-to-noise ratio (SNR) but does not
characterize brain functions at an individual level, a prerequisite for the
application of brain imaging methods in individual diagnostics (cf.
\cite{dubois2016building, eickhoff2020towards}).

% open question
In this study, we investigate if increased hemodynamic activity in the PPA that
is usually revealed by contrasting blocks of pictures can also be revealed by
contrasting selected stimulus features that incidentally occur under more
ecologically conditions.
% we use 2 naturalistic stimuli
To answer this question, we operationalized the perception of both visual and
auditory spatial information using two naturalistic stimuli (see
\citep{hamilton2018revolution, hasson2008neurocinematics,
sonkusare2019naturalistic} for reviews).
% AV stimulus
The operationalization of visual perception is based on an annotation of cuts
and depicted locations in the audio-visual movie ``Forrest Gump'' (AV stimulus).
% AO stimulus
The operationalization of auditory perception is based on an annotation of
speech occurring in the movie's audio-description (AD stimulus).
% GLM contrasts
We then created several general linear model (GLM) contrast to analyze the
corresponding publicly available fMRI datasets
(\href{http://www.studyforrest.org}{studyforrest.org}) of each naturalistic
stimulus \citep{hanke2016simultaneous, hanke2014audiomovie}.

% hypo a
We hypothesized that whole-brain analyses would reveal increased hemodynamic
activity spatially constrained to the PPA as functionally defined by a
traditional visual localizer experiment done with the same subjects
\citep{sengupta2016extension}.
% hypo b: individuals
Further, we hypothesized that an exclusively auditory stimulus could localize
``visual area'' in individual subjects.
% application in individuals
An auditory stimulus substituting a visual experiment would be more practicable
to asses brain functions in visually impaired persons.
% results: group
On a group average level,
% operationalization in naturalistic stimuli
results demonstrate that our approach to operationalize spatial perception
during two naturalistic stimuli lead to increased hemodynamic activity in the
PPA functionally defined by the visual localizer.
% anterior vs. posterior
On both a group and individual subject level, the comparison of results from the
three datasets also suggests functional differences between the anterior and
posterior PPA.
% what does it mean
The current data add evidence to the hypothesis that the PPA might consist of
submodules that work together during visual perception of landscapes (cf.
\citep{baldassano2013differential}).
%
Lastly, our results suggest that the PPA's anterior part can be localized in
individual subjects by contrasting high-level semantic information occurring in
an auditory naturalistic stimulus.

>>>>

% brain mapping via fMRI
Humans can perceive and make sense of their daily environment casually and
seemingly without any effort.
%
Studies in the field of neuropsychology and neuroimaging (e.g.
\citep{penfield1950cerebral, fox1984noninvasive}) have shown that different
parts of the brain are specialized for different perceptual and cognitive
functions.
% occipital cortex for vision -> two pathways
The occipital cortex is considered to be primarily involved in the early stages
of visual perception and giving rise to two distinct pathways that process
different kind of visual information:
% dorsal vs ventral pathway
a dorsal stream (the ``where pathway'') that leads into the parietal lobe, and a
ventral stream (the ``what pathway'') that leads into into the temporal lobe
\citep{goodale1992separate, mishkin1982contribution}.
% PPA
A classic example of a higher-level visual area in the ventral pathway is the
``parahippocampal place area'' (PPA) \citep{epstein1998ppa,
epstein1999parahippocampal}.
% anatomical location != functional location
The PPA is located in the posterior parahippocampal gyrus including adjacent
regions of the fusiform gyrus and anterior lingual gyrus
\citep{epstein2008parahippocampal}.
% neural correlate of scene perception
Increased hemodynamic activity is observed in the PPA when subjects view
pictures of landscapes, buildings or landmarks (compared to e.g. pictures of
faces or tools) during blood oxygenation level-dependent functional magnetic
resonance imaging (BOLD fMRI) \citep{aguirre1998area, epstein2014neural,
epstein1998ppa, troiani2012object}.

% literature review overview 1: imagination & haptic exploration
Increased hemodynamic activity in the PPA generalizes to mental imagination of
landscapes \citep{ocraven2000mental} and haptic exploration of scenes
constructed from LEGO blocks \citep{wolbers2011modality}.
% O'Craven: watching pictures
In a study conducted by \cite{ocraven2000mental} participants viewed alternating
blocks of pictures showing familiar places and famous faces during an initial
experimental paradigm.
% O'Craven: mental imagery
In a subsequent paradigm, participants were instructed to ``form a vivid mental
image'' of the previously viewed pictures.
% O'Craven results
The PPA showed increased activation during imagination of places compared to
faces but the imagination tasks showed a smaller activation level compared to
the perceptual task.
% Wolbers: haptic exploration
In a block design study conducted by \cite{wolbers2011modality} the PPA of
[sighted] participants showed increased activation during a delayed
match-to-sample task of haptically explored scenes constructed from LEGO bricks
compared to abstract geometric objects.
% Wolbers: connectivity analysis

% Aziz (2008): place related sentences
To our knowledge only one study \citep{aziz2008modulation} compared hemodynamic
activity levels in the PPA that were correlated with different semantic
categories occurring in spoken language.
% Aziz' stimuli; ``The Taj Mahal faces a long thin reflecting pool'', ``Marilyn
% Monroe has a large square jaw'', ``The television has a long antenna''
\cite{aziz2008modulation} used sentences that described famous or generic places
and faces, or objects.
% Aziz' tasks:
Participants were instructed to press a button whenever the sentence described
an inaccurate or improbable fact (e.g. ``Marilyn Monroe has a large square
jaw'').
% Aziz' results:
Activation in the left, but not right, PPA was significantly reduced when
participants listened to place-related sentences compared to listening to
face-related sentences. Moreover, this effect was only observed in sentences
involving famous places.

% summary if literature review
In summary, studies suggest that the PPA does preferably but not exclusively
respond to visually presented spatial information.
% three things in common
All reviewed studies share three common aspects:
% designed stimulus set
First, they employed a small sets of carefully chosen and conceptualized
stimuli.
% block-design & task
Second, studies used block-design paradigms to maximize detection power.
Further, they usually employed an explicit (perceptual) judgement task.
% conceptualized stimuli and block: disadvantages
Block-design studies that use conceptualized stimuli and a task have the
advantage of controlling confounding variables (e.g. color, luminance, size,
spatial frequencies, sentence length), maximizing detection power, and keeping
participants paying attention to the stimuli.
% conceptualized stimuli and block: disadvantages
Nevertheless, small sets of conceptualized stimuli and block-design paradigms
lack external and ecological validity \citep{westfall2016fixing,
hasson2004intersubject} because they poorly resemble how we, free of a[n
explicit] perceptual task, experience our rich/multidimensional and continuous
environment that our brains are accommodated to
\citep{sonkusare2019naturalistic}.
% Moreover, study participants can and will become aware about which perceptual
% process is examined in the study by merely being exposed to the experimental
% paradigm.
Third, the reviewed studies averaged results across 7-11 subjects.
% averaging pros and cons
The averaging approach improves signal-to-noise ratio (SNR) but does not
characterize brain functions at an individual level, a prerequisite for the
application of brain imaging methods in individual diagnostics (cf.
\cite{dubois2016building, eickhoff2020towards}).

% open question
In this study, we investigate if increased hemodynamic activity in the PPA that
is usually revealed by contrasting blocks of pictures can also be revealed under
more ecologically conditions.
% we use 2 naturalistic stimuli
To answer this question, we operationalized the perception of both visual and
auditory spatial information using two naturalistic stimuli (see
\citep{hamilton2018revolution, hasson2008neurocinematics,
sonkusare2019naturalistic} for reviews).
% AV stimulus
The operationalization of visual perception is based on an annotation of cuts
and depicted locations in the audio-visual movie ``Forrest Gump'' (AV stimulus).
% AO stimulus
The operationalization of auditory perception is based on an annotation of
speech occurring in the movie's audio-description (AD stimulus).
% GLM contrasts
We then created several general linear model (GLM) contrast to analyze the
corresponding publicly available fMRI datasets of each naturalistic stimulus
\citep{hanke2016simultaneous, hanke2014audiomovie}
(\href{http://www.studyforrest.org}{studyforrest.org}).

% hypo a
We hypothesized that whole-brain analyses would reveal increased hemodynamic
activity spatially constrained to the PPA as functionally defined by a
traditional visual localizer experiment done with the same subjects
\citep{sengupta2016extension}.
% hypo b: individuals
Further, we hypothesized that an exclusively auditory stimulus could localize a
``visual area'' in individual subjects.
% application in individuals
An auditory stimulus substituting a visual experiment would be more practicable
to asses brain functions in visually impaired persons.
% results: group
On a group average level,
% operationalization in naturalistic stimuli
results demonstrate that our approach to operationalize spatial perception
during two naturalistic stimuli lead to increased hemodynamic activity in the
PPA.
% anterior vs. posterior
On both a group and individual subject level, the comparison of results from the
three datasets also suggests functional differences between the anterior and
posterior PPA.
% what does it mean
The current data add evidence to the hypothesis that the PPA might consist of
submodules that work together during visual perception of landscapes (cf.
\citep{baldassano2013differential}).
%
Lastly, our results suggest that the PPA's anterior part can be localized in
individual subjects by contrasting high-level semantic information occurring in
an auditory naturalistic stimulus.
